\SW: II. New lens candidates from the CFHTLS discovered through citizen science

### 2015-04-10 doc/sw-cfhtls.tex
read through and fixed text in Sec 3-5

We begin by considering all elliptical galaxies at $z<1$ in our parent \cfhtls
catalog as potential lens candidates for the simulated sample. To avoid using a
known lens galaxy for our simulation purpose, we exclude all those galaxies
whose positions match with the lensing galaxies from the SARCS
samples within 2~arcsec (XXX check).

For each galaxy, the average number of source objects (either quasars or
galaxies) above a minimum luminosity $L_{\rm min}$ in the background that may get lensed
can be calculated as
\be
\label{eqn:nsrc}
N_{\rm src} = \int_{z_l}^\infty n_{\rm src}(>L_{\rm min},z_s)  \sigma_{\rm
lens}(\sigma_v,z_l,z_s,q) \frac{\rm d V}{{\rm d}z_s}
{\rm d}z_s
\ee
where
\be
\label{eqn:nlum}
n_{\rm src}(>L_{min},z_s)= \int_{L_{\rm min}}^\infty \Phi(L',z_s) {\rm d}L'
\ee
Here, $\Phi(L',z_s)$ denotes the source luminosity function per unit comoving
volume, $\sigma_{\rm lens}$ denotes the angular lens cross-section, which depends upon
the lens redshift ($z_l$), source redshift ($z_s$), the lens velocity
dispersion $\sigma_v$ as well as the projected axis ratio of the lens
ellipticity, $q$.

In order to calculate the lensing cross-section, we first calculate the
luminosity of each potential lensing galaxy using the photometric redshifts
($z_l$) from the LRG catalog. We use the $L-\sigma$ scaling relation from the
bright sample of \citep{Parker2005} to set the velocity dispersion of the halo
hosting the galaxy which will be later used in the model
\be
\label{magstar2}
\sigma_v=142 \left(\frac{L}{L_{*}}\right)^{1/3} \,.
\ee
We assume that the knee of the luminosity function of galaxies, $L_*$ evolves
such that there is a decline of $1.5$ magnitudes between $z=1$ to $z=0$
\citep{Faber2007}.\footnote{We anchor our $L_*$ evolution at low redshifts using the
determination of $L_*$ in the r-band by \citet{Blanton2001}. To maintain
consistency in magnitude systems, we have converted the CFHT MegaCAM magnitudes
to SDSS magnitudes and k-corrected them to $z=0.1$.}

%first calculate the
%luminosity and velocity dispersion of each potential lensing galaxy using the
%CFHT Megacam $g$ and $r$ band magnitudes along with the photometric redshift
%($z_l$) from the LRG catalog.  The Megacam magnitudes are converted to SDSS
%magnitudes\footnote{
%http://www3.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/megapipe/docs/filters.html} and are
%further k-corrected to redshift $z=0.1$ \citep{Frei1994}. We assume that the
%evolution of galaxy luminosities is similar to that determined by
%\citep{Faber2007}, that is, a decline of $1.5$ in the $m_{r*}$ from redshift
%$z=1$ to $z=0$ (see Eq.~\ref{magstar}).

%\be
%\frac{L}{L_{*}}=10.0^{-0.4~(m_{r \rm SDSS}-m_{r*})}
%\ee
%
%where $m_{r*}$ is
%\be
%\label{magstar}
%m_{r*}=-20.44+ 1.5~(z_{l}-0.1) \,.
%\ee
%
%We use the $L-\sigma$ relation from
%\citep{Parker2005} to get the velocity dispersion as given in Eq.~\ref{magstar2}.

%\be
%\label{magstar2}
%\sigma=142 \left(\frac{L}{L_{*}}\right)^{1/3}
%\ee

We assume a singular isothermal ellipsoid model for each of our galaxies
\citep{Kormann1994}, such that the convergence is given by
\be
\kappa (x,y) = \frac{b \sqrt{q}}{2}\frac{1}{\left( \theta_1^2 + q^2\theta_2^2\right)^2}
\ee
Here b is called the Einstein radius, and its dependence on the velocity
dispersion of the SIE is given by
\be
b = 4\pi\,
\left(\frac{\sigma_v^2}{c^2}\right)\left(\frac{D_{ls}}{D_{s}}\right) \,.
\ee
The SIE model results in a caustic and a pseudo-caustic on the source plane:
which demarcate the regions of different image multiplicities. Parametric
solutions, $r(\theta)$ (where $\theta$ is the polar angle), for the caustics in
such a model were given by \citet{Keeton2000b}. We use these parametric
solutions to obtain the area of the lensing cross-section, $\sigma$, for every
galaxy,
\be
\sigma_{\rm lens}=\frac{b^2 q}{2} \, \int_0^{2\pi} r^2(\theta) d\theta\,.
\ee
At every polar angle we use the maximum of the radial and tangential caustic for
the purpose of calculating the cross-section.

%where
%\be
%b_I =  \sqrt{q}\, 4\pi\,
%\left(\frac{\sigma_v^2}{c^2}\right)\left(\frac{D_{ls}}{D_{s}}\right)\,.
%\ee
%Here $D_s$ and $D_{ls}$ are angular diameter distances to
%the source, and between the lens and source, respectively.

%Next, if the foreground galaxy can act as a lens and has at least one source in the
%background, then we determine a redshift ($z_s$) and $i$-band magnitude of the
%background source(s). We assume two types of background sources namely, galaxies
%and quasars. For each source, the redshift and magnitude are generated by drawing randomly from
%the following redshift and luminosity distributions. For
%galaxies, we assume the redshift distribution is

We use the results of \citet{Faure2009} to specify the source luminosity
function of galaxies. We assume that the redshift distribution of sources is
given by
\be
\label{eqn:ps}
p_s=\frac{\beta z_s^2 {\rm exp}({\frac{z_s}{z_0(m_{\rm lim})}})^\beta}{\Gamma(3/\beta)z_0^3(m_{\rm lim})}
\ee
where $\beta=3/2$ and $z_0(m_{\rm lim})=0.13m_{\rm lim} - 2.2$ and the source
counts as a function of the limiting magnitude are given by
\be
\label{eqn:ns}
n_s=\int^{m_{\rm lim}}_{-\infty} \frac{n_0 {\rm d}m}{\sqrt{10^{2a(m_1-m)}+10^{2b(m_1-m)}}}
\,,
\ee
with parameters $a=0.30$, $b=0.56$, $m_1=20$ and $n_0=3\times10^3~deg^{-2}$.

For quasars, we assume the luminosity function prescription of \citep{Oguri2010}
and adopt k-corrections by \citep{Richards2006}.
%alp=-0.5;
%kcorr=-2.5*(1 + alp)*log10(1+zz);
%Dlum=cc.Dlofz(zz)/p.hval;
%DM=5*log10(Dlum);
%Mabs=mag-kcorr-DM-25.0;
The luminosity function is expressed as
\be
\frac{{\rm d}\Phi}{{\rm d}M}=\frac{\Phi_{*}}{10^{0.4(\alpha+1)(M_{\rm abs}-M_{*})} + 10^{0.4(\beta+1)(M_{\rm abs}-M_{*})} }
\ee
where the normalization, $\phi_{*}=5.34\times10^{-6} h^3$ Mpc$^{-3}$ and break
magnitude, $M_*=-20.90 + 5 {\rm log} h - 2.5 {\rm log} f(z)$. The redshift
dependent factor in $M_*$ is given by
\be
f(z)=\frac{e^{\zeta z_s}(1+e^{\xi z_*})}{(\sqrt{e^{\xi z_s}}+\sqrt{e^{\xi z_*}})^2} \,.
\ee
We adopt the best-fit values $\zeta=2.98$, $\xi=4.05$, $z_{*}=1.60$
\citep{Oguri2010}. For the faint end slope, we use $\beta=-1.45$ whereas for
the bright end slope, we use $\alpha=-3.31$ when $z_s<3$ and $\alpha=-2.58$ at
higher redshifts, as prescribed by \citep{Oguri2010}.
respectively.

With the cross-section, and the luminosity functions specified, we calculate the
expected number of sources behind a candidate galaxy using Equation
\ref{eqn:nsrc}. We need to generate a large number of simulated lenses (larger
than the number of real galaxy lenses we expect to find in CFHTLS) in order to
use it as a training sample as well as for calibrating the performance of our
citizen scientists (see Paper I). Therefore, we artificially boost the average
number of sources by a factor (see Table~\ref{tab:thresh}, which increases the
occurrence of lensing. We draw a Poisson deviate with a mean equal to the
boosted average number of sources. If the Poisson deviate is greater than zero,
then we mark this galaxy to act as a lens galaxy. We then draw the properties of
the lensed source based on the redshift and the luminosity distribution of the
sources. The source positions with respect to the lens are drawn randomly from
an area inside the caustic.

Next, we determine properties of the background source for every lens. We follow
similar procedures for both background galaxies and quasars. We perform
ray-tracing for all of the $N_{\rm src}$ sources using the publicly available
code \gravlens \citep{Keeton2000} and choose sources that satisfy our
selection criteria. We determine fluxes of the lensed images and the total
magnification of each of the lensed source. We draw a random source for which
the flux of the second brightest lensed image and the total magnification of all
lensed images are above the thresholds given in Table~\ref{tab:thresh}.

Since we want to produce realistic looking lens systems, we simulate lenses in
each of the five \cfhtls~filters. The colors of the background source galaxies are drawn
randomly from the photometric CFHTLenS catalog
\citep{Hildebrandt2012,Erben2013}.  Similarly, we use a quasar catalog from the
SDSS Data Release 9 \citep{Paris2012} from which colors are drawn to simulate
quasar lenses. Next, we assume deVaucoleur's profile to account for the size
and shape of the galaxies. The ellipticity and the position angle (PA) are drawn
randomly between the range given in Table~\ref{tab:thresh}. The effective
radius of the galaxy is estimated from the Luminosity$-$size relation
\citep{Bernardi2003} (with a redshift scaling to account for some size
evolution) given by
\be
R_{\rm eff}= 10^{0.52} \frac{L_r^{2/3}}{{(1+z_s)}^2}
\ee
where $L_r=L_s/10^{10.2}$. On the other hand, quasars are assumed to follow a
Gaussian profile where the $\sigma$ is equated to that of the median seeing for
every filter. The median seeing values are taken from Table 4 of the official
Terapix T0007 release explanatory document \footnote{
    http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}.

@@

At group or cluster-scales, the mass distribution is more complex. The
convergence in the inner regions, which are typically responsible for the
multiple lensed images, arises from not only the brightest group galaxy (BGG) at
the center, but also has contribution from the dark matter component as well as
the satellite galaxies \citep{Oguri2005,Oguri2006}. We generate a basic group
catalog based on the magnitudes and photometric redshifts available for the
\cfhtls. We select all galaxies with 10$^{10.8} M_\odot$ as plausible BGGs. We
select the member galaxies such that their photometric redshifts are within
$\delta z = 0.01$ of the BGG and within an aperture of $250$~Kpc. If another BGG
is found with the aperture, then the fainter BGG is removed from our list of
BGGs.

@@

We assume a constant mass-to-light ratio of $3 \times 0.7$~h~$M_{*}/L_{*}$, to
convert the BGG luminosity to a stellar mass estimate. The stellar mass$-$halo
mass relation \citep{Behroozi2013}, including random scatter, is then used to
calculate the halo mass for the lens. We adopt an NFW density profile for the
underlying dark matter halo. Given the halo mass, other key parameters such as
the scale radius ($r_s$) and the density at the scale radius ($\rho_s$) can be
determined for an NFW profile. In addition, we adopt an isothermal ellipsoid
model for the BGG and members whenever the ellipticities are available from the
galaxy catalog (else we use an isothermal sphere).

We calculate the luminosity and velocity dispersion for the BGG and each of the
member galaxies following the same prescription as in
Section~\ref{sect:gallens}. To calculate the average number of sources that get
lensed by such a system, we need to calculate the lensing cross-section for each
of these potential lensing groups. The complexity in the lens models makes it
analytically intractable to calculate the size of the caustics\footnote{The lens
mass distribution determines size and shape of the caustics. Any source
located within the caustics will form multiple lensed images which is the
criteria for strong lensing. To further understand caustics, see XXX.}.
Therefore, we generate the caustics numerically using \gravlens and then
determine the area covered by the caustics. We consider only galaxies as our
background source population since group or cluster-scale quasar lenses are not
expected to be extremely rare in the \cfhtls (check XXX).  Following the same procedure
as described in Section~\ref{sect:gallens}, we calculate the number of galaxies
expected to lie behind every potential lensing group (see Eq.~\ref{eqn:nsrc}).
As before, for each background galaxy within the lens cross-section, a redshift
and an $i$-band magnitude is determined by drawing galaxies randomly from the
respective distributions (see Eqs.~\ref{eqn:ps}-\ref{eqn:ns}).

All those groups that are found to have no background galaxies within the
cross-sectional area are rejected and the rest are included as potential lenses.
As mentioned earlier, we artificially boost the total number of sources behind
every lens but ensure that (check XXX) the statistical properties such as the profile of the
image separation distribution are not affected (see \Fref{fig:remudist}). We
follow the same procedure and apply the same thresholds to determine properties
of the lensed galaxies for every lens as are described for galaxy-galaxy lenses
in the previous section. The simulated lensed images are then added on top of
the real \cfhtls~images with the BGGs as the center.

@@

In this section, we describe some of the properties of our simulated sample for
each of the three types of lens samples.

\Fref{fig:remudist} shows the Einstein radius distribution for the
galaxy-scale (dashed for background quasars and dotted for background galaxies)
and groups-scale simulated lenses. For comparison, we show the expected
distributions (blue dashed-dotted curves) for an SIS-like density profile at
galaxy-scales and an NFW+Hernquist profile at groups-scales. The
theoretical curves are taken from \citep{More2012} wherein the models are
explained in detail. We note that the model we adopt at groups-scale also
includes SIS or SIE components for the group members unlike the theoretical
prediction. The theoretical curves have arbitrary normalizations.

We show the redshift, magnitude and ellipticity distributions of the lensing
galaxies in the three panels of \Fref{fig:lensprop}, respectively. We also show
the distributions of respective properties of the SARCS lenses from
\citep{More2012} for comparison with arbitrary normalizations. We note the
qualitative similarities between the simulated and the real lens samples.

%Similarly, we show the ellipticity and the position angle of the simulated lens
%galaxy population extracted from the T0007 release of the \cfhtls~catalogs in the
%left and right panels of \Fref{fig:lensprop}, respectively. As before, the
%dashed-dotted (blue) curves show the same distribution for the SARCS lens
%population with arbitrary normalization \citep{More2012}.

%We show the source redshift and magnitude distributions for each of the three
%lens samples in the left and right panels of \Fref{fig:szmdist},
%respectively. The peak of the source redshift distribution both for the galaxies
%and quasars is known to be between $2<z<3$ from various lens samples (add
%references XXX). This consistent with our simulated sample as expected. The
%magnitude distribution of the simulated sample, on the other hand, is
%specifically tailored to the requirements of \sw.  The goal was to generate a
%lens sample that has a fair balance of both bright and faint lensed images since
%the training sample should neither be too easy nor too difficult for the citizen
%scientists.

We produce catalogs with lens and source properties for each of the three types
of lenses. These catalogs are available from
https://github.com/anupreeta27/SIMCT. The catalogs typically have lens
position, redshift, magnitudes, Einstein radius, ellipticity (whenever
available) and shear (for galaxy-scale lenses only). For the background
sources, we provide the offset from the lens center, redshift, magnitudes,
total magnification, number of lensed images. Additionally, when possible,
ellipticity and effective radius of the background galaxies have also been
provided.

@@

\subsection{Limitations of the simulated sample}
The simulated lens sample, although realistic, is not perfect, due to the
simplicity of the lensing models and our limited understanding of the
uncertainties in the parameters for the models. Comments from citizen
scientists were very helpful in order to sort some of these failures. Here, we
describe some of the cases or aspects in which the simulations were known to
have failed (or appeared unrealistic). This typically happens for a small
fraction of the simulated images (roughly 5 percent of
the cases)\footnote{Based on the hashtag simfail in the discussion forum
TALK.}.

The parameters required by various scaling relations and the models primarily
depend on the photometry of the galaxies, groups and quasars detected in the
survey. For galaxy-scale lenses, the lensing galaxies at higher redshifts or
which are fainter have poor photometric redshift measurements. Consequently
these galaxies are assigned a wrong luminosity and velocity dispersion
estimates, which sometimes result in simulated lenses which look implausible.
For example, the lensed images for some of the failed simulations have larger
image separation than expected given the visual priors from the galaxy.

At group-scales, the photometric and redshift estimates are used when define
the group membership. Therefore, errors in redshift estimates generate galaxy
groups with BGG or member galaxies with dissimilar properties. In some cases,
low redshift spiral galaxies have been incorrectly assigned high redshift.
Spiral galaxies are typically less massive and low redshift spiral galaxies are
unlikely to act as gravitational lenses. Hence, some of the simulated lenses
were not as convincing.

We also use a single Sersic component to describe the light profiles of
background galaxies. This is clearly not the most accurate description for
galaxies, especially, star-forming galaxies which form a significant fraction
of the lensed galaxy population. Star-forming galaxies have complex structures
such as star forming knots, spiral arms, bars and disks. The simulated lensed
images do not display these features.

@@

Citizen scientists not only need training to identify gravitational lenses, but
also to reject images which either contain impostors or have no lenses. Hence,
in addition to the simulated lenses, we added a sample of duds and false
positives to the training sample. Duds are images which have been visually
inspected by experts and confirmed to contain no lenses.  False positives (FPs)
are systems which look like lenses but are not, for example, spiral galaxies,
starforming galaxies, chance alignments of features arranged in a lensing
configuration and stars.

We selected a sample of 450 duds for the \StageOne classification in \sw
and a sample of 500 false positives for the \StageTwo inspection. The
sample of false positives was selected from the candidates which passed
the \StageOne of \sw. We note that this is the first time, we have a
systematically compiled sample of visually inspected false positives by
the \sw volunteers and categorized by the science team. The data products are
available at http://spacewarps.org/\#/projects/CFHTLS/. Such a sample is
tremendously helpful for training and understanding performances of
various lens finding algorithms \citep{Chan2014}.

@@

\section{Methodology to produce the \sw-\cfhtls Lens Sample}
\label{sec:swap}
The \sw works as a single unified neural network which uses the method of visual
inspection to find gravitational lenses. The volunteers are shown images at two
stages. At \StageOne, volunteers are requested to undertake a rapid inspection to
select lens candiates ranging from possible lenses to almost certain lenses. At
\StageTwo, volunteers carefully inspect the candidates from \StageOne and select only
promising lens candidates. A daily snapshot of classifications performed by
volunteers are provided to the science team every night. This daily batch is
analyzed by the Space Warps Analysis Pipeline (SWAP). The philosophy and the
details of SWAP are described in detail in Paper I, here we briefly summarize
how it works.

>>>>

We begin by considering all elliptical galaxies at $z<1$ in our parent
\cfhtls catalog \citep[(XXX confirm with Raphael)]{Gavazzi2014} as
potential lens candidates for the simulated sample. To avoid using a
known lens galaxy for our simulation purpose, we exclude all those
galaxies whose positions match with the lensing galaxies from the SARCS
samples within 2~arcsec (XXX check).

For each galaxy, the average number of source objects (either quasars or
galaxies) above a minimum luminosity $L_{\rm min}$ in the background that may get lensed
can be calculated as
\be
\label{eqn:nsrc}
N_{\rm src} = \int_{z_l}^\infty n_{\rm src}(>L_{\rm min},z_s)  \sigma_{\rm
lens}(\sigma_v,z_l,z_s,q) \frac{\rm d V}{{\rm d}z_s}
{\rm d}z_s
\ee
where
\be
\label{eqn:nlum}
n_{\rm src}(>L_{\rm min},z_s)= \int_{L_{\rm min}}^\infty \Phi(L',z_s) {\rm d}L' \,.
\ee

Here, $\Phi(L',z_s)$ denotes the source luminosity function per unit comoving
volume, $\sigma_{\rm lens}$ denotes the angular lens cross-section, which depends upon
the lens redshift ($z_l$), source redshift ($z_s$), the lens velocity
dispersion $\sigma_v$ as well as the projected axis ratio of the lens
ellipticity, $q$.

In order to calculate the lensing cross-section, we first calculate the
luminosity of each potential lensing galaxy using the photometric redshifts
($z_l$) from the parent galaxy catalog. Next, we use the $L-\sigma$ scaling relation from the
bright sample of \citep{Parker2005} to set the velocity dispersion of the halo
hosting the galaxy which will be later used in the model
\be
\label{magstar2}
\sigma_v=142 \left(\frac{L}{L_{*}}\right)^{1/3} \,.
\ee
We assume that the knee of the luminosity function of galaxies, $L_*$ evolves
such that there is a decline of $1.5$ magnitudes between $z=1$ to $z=0$
\citep{Faber2007}.\footnote{We anchor our $L_*$ evolution at low redshifts using the
determination of $L_*$ in the r-band by \citet{Blanton2001}. To maintain
consistency in magnitude systems, we have converted the CFHT MegaCAM magnitudes
to SDSS magnitudes and k-corrected them to $z=0.1$.}

We adopt a singular isothermal ellipsoid (SIE) model for each of our galaxies
\citep{Kormann1994}, such that the convergence is given by
\be
\kappa (x,y) = \frac{b \sqrt{q}}{2}\frac{1}{\left( \theta_1^2 + q^2\theta_2^2\right)^2} \,.
\ee
Here, $b$ is called the Einstein radius, and its dependence on the velocity
dispersion of the SIE is given by
\be
b = 4\pi\,
\left(\frac{\sigma_v^2}{c^2}\right)\left(\frac{D_{ls}}{D_{s}}\right) \,.
\ee
The SIE model results in a caustic and a pseudo-caustic on the source
plane: which demarcate the regions of different image multiplicities. We
make use of the parametric solutions, $r(\theta)$, for the caustics in
such a model from \citet{Keeton2000b} where $\theta$ is the polar angle.
We take the maximum of the radial and tangential caustic at every polar
angle in order to obtain the area of the lensing cross-section,
$\sigma_{\rm lens}$, for every galaxy,
\be
\sigma_{\rm lens}=\frac{b^2 q}{2} \, \int_0^{2\pi} r^2(\theta) d\theta \,.
\ee

The luminosify functions of the background galaxies and quasars are
determined as follows. We use the results of \citet{Faure2009} to specify the luminosity
function of galaxies where the redshift distribution of sources is
given by
\be
\label{eqn:ps}
p_s=\frac{\beta z_s^2 {\rm exp}({\frac{z_s}{z_0(m_{\rm lim})}})^\beta}{\Gamma(3/\beta)z_0^3(m_{\rm lim})}
\ee
where $\beta=3/2$ and $z_0(m_{\rm lim})=0.13~m_{\rm lim} - 2.2$ and the source
counts as a function of the limiting magnitude are given by
\be
\label{eqn:ns}
n_s=\int^{m_{\rm lim}}_{-\infty} \frac{n_0 {\rm d}m}{\sqrt{10^{2a(m_1-m)}+10^{2b(m_1-m)}}}
\,,
\ee
with parameters $a=0.30$, $b=0.56$, $m_1=20$ and $n_0=3\times10^3~deg^{-2}$.

For quasars, we assume the luminosity function prescription of \citep{Oguri2010}
and adopt k-corrections by \citep{Richards2006}.

The luminosity function is expressed as
\be
\frac{{\rm d}\Phi}{{\rm d}M}=\frac{\Phi_{*}}{10^{0.4(\alpha+1)(M_{\rm abs}-M_{*})} + 10^{0.4(\beta+1)(M_{\rm abs}-M_{*})} }
\ee
where the normalization, $\phi_{*}=5.34\times10^{-6} h^3$ Mpc$^{-3}$ and break
magnitude, $M_*=-20.90 + 5 {\rm log} h - 2.5 {\rm log} f(z)$. The redshift
dependent factor in $M_*$ is given by
\be
f(z)=\frac{e^{\zeta z_s}(1+e^{\xi z_*})}{(\sqrt{e^{\xi z_s}}+\sqrt{e^{\xi z_*}})^2} \,.
\ee
We adopt the best-fit values $\zeta=2.98$, $\xi=4.05$, $z_{*}=1.60$
\citep{Oguri2010}. For the faint end slope, we use $\beta=-1.45$ whereas for
the bright end slope, we use $\alpha=-3.31$ when $z_s<3$ and $\alpha=-2.58$ at
higher redshifts, as prescribed by \citep{Oguri2010}.

With the cross-section, and the luminosity functions specified, we
calculate the expected number of sources behind a candidate lensing
galaxy using Eq.~\ref{eqn:nsrc}. We need to generate a large number of
simulated lenses (larger than the number of real galaxy lenses we expect
to find in \cfhtls) in order to have a reasonably large and diverse
training sample for thousands of \sw volunteers. Therefore, we
artificially boost the average number of sources by a factor (see
Table~\ref{tab:thresh}, which increases the occurrence of lensing. We
draw a Poisson deviate with a mean equal to the boosted average number
of sources. If the Poisson deviate is greater than zero, then 
this galaxy is flagged as a potential lensing galaxy with
Nsrc number of background sources. {\bf XXXX check if this is correct.}

Next, we determine properties of the background sources for every lens.
We follow similar procedures for both background galaxies and quasars.
We draw source redshifts and luminosities from the aforementioned
distributions. The source positions with respect to the lens are drawn
randomly from an area inside the caustic.  We perform ray-tracing for
all of the $N_{\rm src}$ sources using the publicly available code
\gravlens \citep{Keeton2000} and choose sources that satisfy our
selection criteria given below. We determine fluxes of the lensed images and the
total magnification of each of the lensed source. We draw a
source randomly for which the flux of the second brightest lensed image and the
total magnification of all lensed images meet the thresholds given
in Table~\ref{tab:thresh}.

Since we want to produce realistic looking lens systems, we simulate lenses in
each of the five \cfhtls~filters. The colors of the background galaxies are drawn
randomly from the photometric CFHTLenS catalog
\citep{Hildebrandt2012,Erben2013}.  Similarly, we use a quasar catalog from the
SDSS Data Release 9 \citep{Paris2012} from which colors are drawn to simulate
quasar lenses.  Next, we assume deVaucoleur's profile to account for the size
and shape of the galaxies. The ellipticity and the position angle (PA) are drawn
randomly from within the range given in Table~\ref{tab:thresh}. The effective
radius of the galaxy is estimated from the Luminosity$-$size relation
\citep{Bernardi2003} (with a redshift scaling, to account for size
evolution) given by
\be
R_{\rm eff}= 10^{0.52} \frac{L_r^{2/3}}{{(1+z_s)}^2}
\ee
where $L_r=L_s/10^{10.2}$. On the other hand, quasars are assumed to follow a
Gaussian profile where the full-width-at-half-maximum is equated to that of the median seeing for
every filter. The median seeing values are taken from Table 4 of the official
Terapix T0007 release explanatory document \footnote{
    http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}.

@@

At group or cluster-scales, the mass distribution is more complex. The
convergence in the inner regions, which are typically responsible for the
multiple lensed images, arises from not only the brightest group galaxy (BGG) at
the center, but also from the dark matter component and
the satellite galaxies \citep{Oguri2005,Oguri2006}. We generate a basic group
catalog based on the magnitudes and photometric redshifts available for the
\cfhtls. We select all galaxies with 10$^{10.8} M_\odot$ as plausible BGGs. We
select the member galaxies such that their photometric redshifts are within
$\delta z = 0.01$ of the BGG and within an aperture of $250$~Kpc. If another BGG
is found with the aperture, then the fainter BGG is removed from our list of
BGGs.

@@

We assume a constant mass-to-light ratio of $3 \times
0.7$~h~$M_{*}/L_{*}$, to convert the BGG luminosity to a stellar mass
estimate. The stellar mass$-$halo mass relation \citep{Behroozi2013},
including random scatter, is then used to calculate the halo mass for
the lens. We adopt an NFW \citep{Navarro1997} density profile for the
underlying dark matter halo. Given the halo mass, other key parameters
such as the scale radius ($r_s$) and the density at the scale radius
($\rho_s$) can be determined for an NFW profile. In addition, we adopt
an SIE model for the BGG and members whenever the ellipticities are
available from the galaxy catalog (else we use an isothermal sphere,
SIS).

We calculate the luminosity and velocity dispersion for the BGG and each of the
member galaxies following the same prescription as in
\Sref{sect:gallens}. To calculate the average number of sources that get
lensed by such a system, we need to calculate the lensing cross-section for each
of these potential lensing groups. The complexity in the lens models makes it
analytically intractable to calculate the size of the caustics\footnote{The lens
mass distribution determines size and shape of the caustics. Any source
located within the caustics will form multiple lensed images which is the
criteria for strong lensing. To further understand caustics, see XXX.}.
Therefore, we generate the caustics numerically using \gravlens and then
determine the area covered by the caustics. We consider only galaxies as our
background source population since group or cluster-scale quasar lenses are
expected to be extremely rare in the \cfhtls.  Following the same procedure
as described in \Sref{sect:gallens}, we calculate the number of galaxies
expected to lie behind every potential lensing group (see Eq.~\ref{eqn:nsrc}).
As before, for each background galaxy within the lens cross-section, a redshift
and an $i$-band magnitude is determined by drawing galaxies randomly from the
respective distributions (see Eqs.~\ref{eqn:ps}-\ref{eqn:ns}).

All those groups that are found to have no background galaxies within the
cross-sectional area are rejected and the rest are included as potential lenses.
As mentioned earlier, we artificially boost the total number of sources behind
every lens but ensure that (check XXX) the statistical properties such as the profile of the
image separation distribution are not affected (see \Fref{fig:remudist}). We
follow the same procedure and apply thresholds to determine properties
of the lensed galaxies for every lens as are described for galaxy-galaxy lenses
in the previous section. The thresholds are same as those used for
galaxy lenses (see \Tref{tab:thresh}) and are reported within ``()'', if
different for group-scales. The simulated lensed images are then added
to the real \cfhtls~images with the BGGs as the center by following
exactly the same procedure as described in the previous section.

@@

In this section, we describe some of the properties of our simulated sample for
each of the three types of lens samples. We have made an attempt to
generate as realistic a lens sample as possible within the requirements
of the \sw project. The statistical properties of the lens sample are
expected to be similar to real lens samples.

In \Fref{fig:remudist}, we show the Einstein radius distribution for the
galaxy-scale (dashed for background quasars and dotted for background galaxies)
and groups-scale simulated lenses. For comparison, we give the expected
distributions (blue dashed-dotted curves) for an SIS like density profile at
galaxy-scales and an NFW+Hernquist profile at groups-scales. The
theoretical curves are taken from \citet{More2012} wherein the models are
explained in detail. We note that the model we adopt at groups-scale also
includes SIS or SIE components for the group members unlike the theoretical
prediction. The theoretical curves have arbitrary normalizations. We
also show the distribution of the total magnification for all three
samples.

Next, we consider the redshift, magnitude and ellipticity distributions
of the lensing galaxies as shown in \Fref{fig:lensprop}. For reference,
we also show SARCS lenses from \citet{More2012}, with arbitrary
normalizations. We find that the properties of the foreground lenses in
the simulated and the real lens samples are broadly similar.

We produce catalogs with lens and source properties for each of the
three types of lenses.  The catalogs typically have lens position,
redshift, magnitudes, Einstein radius, ellipticity (whenever available)
and shear (for galaxy-scale lenses only). For the background sources, we
provide the offset from the lens center, redshift, magnitudes, total
magnification, number of lensed images. Additionally, when possible,
ellipticity and effective radius of the background galaxies have also
been provided. These catalogs are available from
https://github.com/anupreeta27/SIMCT and simulated lens sample can be
requested from the authors.

@@

\subsection{Limitations of the simulated sample}
The simulated lens sample, although realistic, is not perfect, due to
the simplicity of the lensing models and our limited understanding of
the uncertainties in the model parameters. Comments from citizen
scientists were very helpful in order to sort some of these failures.
Here, we describe some of the cases or aspects in which the simulations
were known to have failed to look realistic. The fraction of such
simulated sample is small (roughly 5 percent)\footnote{This estimate is
based on the \#simfail from TALK, the discussion forum.}.

The parameters required by various scaling relations and the models
primarily depend on the photometry of the galaxies, groups and quasars
detected in the survey. For galaxy-scale lenses, the lensing galaxies at
higher redshifts or which are fainter have poor photometric redshift
measurements. Consequently these galaxies are assigned a wrong
luminosity and velocity dispersion estimates. This then results in
simulated lenses which look implausible or unrealistic.  For example,
the lensed images for some of the failed simulations have larger image
separation than what you expect from the luminosity and/or size of the
galaxy. We roughly expect mass to follow light so more massive galaxy
typically looks brighter and/or bigger.

At group-scales, the photometric and redshift estimates are used when define
the group membership. Therefore, errors in redshift estimates generate galaxy
groups with BGG or member galaxies with dissimilar properties. In some cases,
low redshift spiral galaxies have been incorrectly assigned high redshift.
Spiral galaxies are typically less massive and low redshift spiral galaxies are
unlikely to act as gravitational lenses. Hence, some such instances
do not appear convincing or the lensed images do not have the expected
configuration or image separations.

We also use a single Sersic component to describe the light profiles of
background galaxies. This is clearly not the most accurate description for
galaxies, especially, star-forming galaxies which form a significant fraction
of the lensed galaxy population. Star-forming galaxies have complex structures
such as star forming knots, spiral arms, bars and disks. The simulated lensed
images do not display these features. This is fine for most cases or
mostly for images taken from ground based telescopes such as the CFHT
but sometimes the shapes of the lensed images can appear symmetric or
perfect, especially, if the images are very bright.

@@

Citizen scientists need training not only to identify gravitational
lenses, but also to reject images which either have no lenses or have
objects which are lens impostors. Hence, in addition to the simulated
lenses, we added a sample of duds and false positives to the training
sample. Duds are images which have been visually inspected by experts
and confirmed to contain no lenses. False positives (FPs) are systems
which look like lenses but are not, for example, spiral galaxies,
starforming galaxies, chance alignments of features arranged in a
lensing configuration and stars.

We selected a sample of 450 duds for the \StageOne classification in \sw
and a sample of 500 false positives for the \StageTwo inspection. The
sample of false positives was selected from the candidates which passed
the \StageOne of \sw. We note that this is the first time, we have a
systematically compiled sample of visually inspected false positives by
the \sw volunteers and categorized by the science team. We produce
additional larger sample of a few thousand FPs by scanning through the
low probability images after the completion of \StageTwo. All of these
data products are available at
http://spacewarps.org/\#/projects/CFHTLS/. Such a sample has tremendous
utility for training and testing of various lens finding algorithms
\citep[e.g.,][]{Chan2014}.

@@

\section{Methodology to produce the \sw-\cfhtls Lens Sample}
\label{sec:swap}
The \sw works as a single unified system which uses the method of visual
inspection to find gravitational lenses. For the first \sw lens search,
the volunteers were shown images at two stages. At \StageOne, volunteers
are requested to undertake a rapid inspection to select lens candiates
ranging from possible lenses to almost certain lenses. At \StageTwo,
volunteers carefully inspect the candidates from \StageOne and select
only promising lens candidates. A daily snapshot of classifications
performed by volunteers are provided to the science team every night.
This daily batch is analyzed by the Space Warps Analysis Pipeline
(SWAP). The philosophy and the details of SWAP are described in detail
in Paper I, here we briefly summarize how it works.
